Titeln på avhandlingens språk

### 2023-05-19 body/part1/chapter2-relatedwork/scope.tex
cleaning up for publication

Before these systems can become widely deployed, however, their real-world performance must be thoroughly understood.
\glspl{CPS} often operate in safety-critical environments, and poor \gls{QoS} in \gls{XR} can easily lead to users experiencing discomfort while using the application.
Accurate evaluation, characterization, and optimization of the performance of these systems is therefore critical to ensure reliability, safety, and user satisfaction.
Additionally, accurate performance evaluation enables efficient resource utilization, and enhances system scalability, in turn leading to lower operation costs and a reduced barrier of entry.

Evaluating the performance of Edge environments involving such applications is challenging, due to complex behaviors emerging from the interaction between computation and network in these systems.
Existing literature has attempted to tackle the performance evaluation of Edge systems through analytical, simulation, and experimental approaches.
However, the nature of Edge Computing environments, which involves tight integration between network and compute, requires vast and often unfeasible interdisciplinary competence.
Existing works have been constrained to narrow their scope by prioritizing the comprehensive modeling of either compute or network, resulting in a potential trade-off that may neglect the other aspect.
As we will see in \cref{sec:state-of-the-art}, this has led to an incomplete understanding of the overall performance of Edge systems and applications in literature.

In this dissertation we introduce and study the applicability of an alternative methodology for Edge performance evaluation based on a practical, experimental approach.
We aim to complement literature by providing a method which bridges the gap in knowledge left by existing approaches.
Or work centers of Edge Computing environments involving latency-sensitive applications such as \gls{XR} and \glspl{CPS}, as these are broadly understood to represent ``killer'' use-cases for the Edge.
We also focus exclusively on Cloudlet Computing-like setups.
Although \gls{MEC} is one of the main enablers for \gls{WCA} applications, we consider such network deployments to be \gls{OOS} due to the complexity and financial investment required to deploy a mobile network.
Likewise, we do not target Fog Computing deployments due to this paradigm being a poor fit for latency-sensitive applications.
Thus, although the methodology and tools introduced are architecture-agnostic and should be directly applicable to other kinds of Edge Computing setups than the one used in this dissertation, further work may be needed to validate our results and conclusions in the context of \gls{MEC} and Fog Computing.

The performance evaluation of Edge-enabled \gls{XR} encompasses a wide range of factors, making the scope of research in this area extensive and complex.
The scope of performance evaluation research for \glspl{CPS} is at least equally extensive.
Attempting to cover both these scopes comprehensively in a single dissertation would be prohibitively challenging.
Therefore, in this work, we approach these challenges by focusing on representative use-cases for these application classes.

Specifically, we employ \emph{step-based task-assistance} \gls{WCA} as a representative use case for \gls{XR}, and \glspl{NCS} for \glspl{CPS}.
We focus on these categories for a couple of reasons.
In the case of \gls{WCA}, step-based task-assistance was among the first categories of \gls{WCA} to be developed~\cite{chen2015early}, and thus represents a more mature application class compared to other \glspl{WCA}.
These applications also feature a linear mode of execution, simplifying emulation, and their step-based nature makes the tasks easier to shorten or extend for experimental purposes.

For \glspl{NCS}, we limit our exploration in this dissertation to simple, small-scale, single-plant systems.
The system used corresponds to the \gls{2D} inverted pendulum, using simple \gls{PD} and \gls{PID} controllers.
We have chosen this system for its relative simplicity as well as its broad use in the field of automatic control as one of the fundamental examples of linear control.
Through these examples, we aim to gain insights into the performance evaluation challenges and opportunities offered by Edge Computing these applications and systems, and contribute to the understanding of the broader fields of \gls{XR} and \gls{CPS} performance evaluation in distributed computing environments.

@@

In the following  \cref{sec:state-of-the-art}, we will review the state of the art in the context of performance evaluation of Edge Computing environments, including analytical, simulation, and experimental approaches.
We will evidence how existing approaches fail to fully capture the intricacies of performance evaluation of latency-sensitive, feedback-loop applications on Edge Computing, and make the case for the need for an alternative approach.
We will further discuss existing experimental approaches to \gls{MAR} and \gls{NCS} research, and discuss their limitations.

Finally, we conclude this chapter in \cref{sec:contributions}, where we will discuss in details the contributions of this dissertation.
This includes a discussion of the key sub-contributions of each included publication.

>>>>

\input{tables/scope}

Before these systems can become widely deployed, however, their real-world performance must be thoroughly understood.
\glspl{CPS} often operate in safety-critical environments, and poor \gls{QoS} in \gls{XR} can easily lead to users experiencing discomfort.
Accurate evaluation, characterization, and optimization of the performance of these systems is therefore critical to ensure reliability, safety, and user satisfaction.
Additionally, accurate performance evaluation enables efficient resource utilization, and enhances system scalability, in turn leading to lower operation costs and a reduced barrier of entry.

Evaluating the performance of Edge environments involving such applications is challenging, due to complex behaviors emerging from the interaction between computation and network.
Existing literature has attempted to tackle the performance evaluation of Edge systems through analytical, simulation, and experimental approaches.
However, the nature of Edge Computing environments, which involves tight integration between network and compute, requires vast and often unfeasible interdisciplinary competence.
Existing works have been constrained to narrow their scope by prioritizing the comprehensive modeling of either compute or network, resulting in a trade-off that may neglect the other aspect.
As I will discuss in \cref{sec:state-of-the-art}, this has led to an incomplete understanding of the overall performance of Edge systems and applications in literature.

In this dissertation I introduce and study the applicability of an alternative methodology for Edge performance evaluation based on a practical, experimental approach.
I aim to complement literature by providing a method which bridges the gap in knowledge left by existing approaches.
My work centers of Edge Computing environments involving latency-sensitive applications such as \gls{XR} and \glspl{CPS}, as these are broadly understood to represent ``killer'' use-cases for the Edge.
I also focus exclusively on Cloudlet Computing-like setups (see \cref{tab:scope}).
Although \gls{MEC} is one of the main enablers for \gls{WCA} applications, I consider such network deployments to be \gls{OOS} due to the complexity and financial investment required to deploy a mobile network.
Likewise, I do not target Fog Computing deployments due to this paradigm being a poor fit for latency-sensitive applications.
Thus, although the methodology and tools introduced are architecture-agnostic and should be directly applicable to other kinds of Edge Computing setups than the one used in this dissertation, further work may be needed to validate my results and conclusions in the context of \gls{MEC} and Fog Computing.

The performance evaluation of Edge-enabled \gls{XR} encompasses a wide range of factors, making the scope of research in this area extensive and complex.
The scope of performance evaluation research for \glspl{CPS} is at least equally extensive.
Attempting to cover both these scopes comprehensively in a single dissertation would be prohibitively challenging.
Therefore, in this work, I approach these challenges by focusing on representative use-cases for these application classes.

Specifically, I employ \emph{step-based task-assistance} \gls{WCA} as a representative use case for \gls{XR}, and \glspl{NCS} for \glspl{CPS}.
I focus on these categories for a couple of reasons.
In the case of \gls{WCA}, step-based task-assistance was among the first categories of \gls{WCA} to be developed~\cite{chen2015early}, and thus represents a more mature application class compared to other \glspl{WCA}.
These applications also feature a linear mode of execution, simplifying emulation, and their step-based nature makes the tasks easier to shorten or extend for experimental purposes.

For \glspl{NCS}, I limit my exploration in this dissertation to simple, small-scale, single-plant systems.
The system used corresponds to the \gls{2D} inverted pendulum, using simple \gls{PD} and \gls{PID} controllers.
I have chosen this system for its relative simplicity as well as its broad use in the field of automatic control as one of the fundamental examples of linear control.
Through these examples, I aim to gain insights into the performance evaluation challenges and opportunities offered by Edge Computing these applications and systems, and contribute to the understanding of the broader fields of \gls{XR} and \gls{CPS} performance evaluation in distributed computing environments.

@@

In \cref{sec:state-of-the-art}, I will review the state of the art in the context of performance evaluation of Edge Computing environments, including analytical, simulation, and experimental approaches.
I will evidence how existing works fail to fully capture the intricacies of performance evaluation of latency-sensitive, feedback-loop applications on Edge Computing, and make the case for the need for an alternative approach.
I will further discuss existing experimental approaches to \gls{MAR} and \gls{NCS} research, and discuss their limitations.

Finally, I conclude this chapter in \cref{sec:contributions}, where I will discuss in details the contributions of this dissertation.
This includes a discussion of the key sub-contributions of each included publication.

### 2023-05-19 body/part1/chapter2-relatedwork/tables/scope.tex
cleaning up for publication

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
% \usepackage[table,xcdraw]{xcolor}
% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
\begin{table}[]
    \centering
    \caption{Visualization of the scope of this dissertation with respect to performance evaluation.
    We focus exclusively on Edge Computing based on Cloudlets~\cite{satyanarayanan2009case} and on experimental performance evaluation of \acs{XR} and \acsp{CPS}.}
    \label{tab:scope}
    \begin{tabular}{@{}cccc@{}}
        \toprule
        & \makecell{Analytical\\Modeling} & {Simulations} & {Experimental}                   \\ \midrule
        Cloudlets & \cellcolor[HTML]{9AFF99}Yes & \cellcolor[HTML]{9AFF99}Yes & \cellcolor[HTML]{9AFF99}{\color[HTML]{333333} Yes} \\
        \acs{MEC} & \cellcolor[HTML]{FFCCC9}No              & \cellcolor[HTML]{FFCCC9}No      & \cellcolor[HTML]{FFCCC9}No                         \\
        Fog & \cellcolor[HTML]{FFCCC9}No              & \cellcolor[HTML]{FFCCC9}No      & \cellcolor[HTML]{FFCCC9}No                         \\
        \acs{XR}  & \cellcolor[HTML]{FFCCC9}No              & \cellcolor[HTML]{FFCCC9}No      & \cellcolor[HTML]{9AFF99}{\color[HTML]{333333} Yes} \\
        \acsp{CPS} & \cellcolor[HTML]{FFCCC9}No              & \cellcolor[HTML]{FFCCC9}No      & \cellcolor[HTML]{9AFF99}{\color[HTML]{333333} Yes} \\ \bottomrule
    \end{tabular}
\end{table}

>>>>

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
% \usepackage[table,xcdraw]{xcolor}
% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
\begin{table}[]
    \centering
    \caption{Visualization of the scope of this dissertation with respect to performance evaluation.
    I focus exclusively on Edge Computing based on Cloudlets~\cite{satyanarayanan2009case} and on experimental performance evaluation of \acs{XR} and \acsp{CPS}.}
    \label{tab:scope}
    \begin{tabular}{@{}cccc@{}}
        \toprule
        & \makecell{Analytical\\Modeling} & {Simulations} & {Experimental}                   \\ \midrule
        Cloudlets & \cellcolor[HTML]{9AFF99}Yes & \cellcolor[HTML]{9AFF99}Yes & \cellcolor[HTML]{9AFF99}{\color[HTML]{333333} Yes} \\
        \acs{MEC} & \cellcolor[HTML]{FFCCC9}No              & \cellcolor[HTML]{FFCCC9}No      & \cellcolor[HTML]{FFCCC9}No                         \\
        Fog & \cellcolor[HTML]{FFCCC9}No              & \cellcolor[HTML]{FFCCC9}No      & \cellcolor[HTML]{FFCCC9}No                         \\
        \acs{XR}  & \cellcolor[HTML]{FFCCC9}No              & \cellcolor[HTML]{FFCCC9}No      & \cellcolor[HTML]{9AFF99}{\color[HTML]{333333} Yes} \\
        \acsp{CPS} & \cellcolor[HTML]{FFCCC9}No              & \cellcolor[HTML]{FFCCC9}No      & \cellcolor[HTML]{9AFF99}{\color[HTML]{333333} Yes} \\ \bottomrule
    \end{tabular}
\end{table}
